---
title: "Ploidy_Analysis"
author: "Lukas P. Grossfurthner"
output:
  github_document:
    toc: yes
    toc_depth: 3
  pdf_document:
    toc: yes
    toc_depth: 3
  html_document:
    toc: yes
    toc_depth: 3
    theme: united
    keep_md: true
date: "`r Sys.Date()`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(readxl)
library(ggrepel)
```

## Ploidy Analysis

Workflow used to estimate genome sizes and to visualize the resulting data. For screening greenhouse samples, samples were run on the FCM once, for the field data, samples were run three times and the average of the measurements was taken.  


```{r read_data, message=F}
genome_size_std <- 0.535 # genome size of A. negundo as identified by Contreras et al. 2007
greenhouse_data <- read_xlsx("data/tables/Atri_greenhouse_ploidy.xlsx", sheet=2)
triploid_offspring_PI <- read_excel("../triploid/greenhouse_samples/data/Triploid_Offspring_measurements.xlsx")
colnames(greenhouse_data) <- c("Sample_ID", "Garden", "Source", "Flat_ID", "Ratio","C2_value", "Original_Reseed")
field_data <- read_xlsx("../triploid/greenhouse_samples/data/Atri_field_ploidy.xlsx")

```

## Estimate genome sizes based on DAPI of samples and combine data sets


```{r read_and_manipulate_data, echo=FALSE}
Oregon_1 <- field_data %>% group_by(Sample) %>% 
  mutate(C2_value_Acne=Sample_mean/Std_mean*(genome_size_std*2)) %>% 
  select(Sample, C2_value_Acne) %>% 
  summarize(mean_C2=mean(C2_value_Acne)) %>% 
  mutate(Std="Acne",
         Sample_ID=substr(Sample, 4, nchar(Sample))) %>% 
  separate_rows(Sample_ID, sep = "_") %>% 
  mutate(Sample_ID= paste0(substr(Sample, 1,3), Sample_ID),
         Site=paste("Oregon", substr(Sample_ID, 2,2))) %>% 
  select(Sample_ID, Site, mean_C2, Std) %>% 
  dplyr::rename("C2_value"=mean_C2)

Greenhouse <- greenhouse_data %>% mutate(Site=Garden, 
                           Sample_ID=paste(Source,Sample_ID, sep="_"),
                           Std="Atca") %>% 
  select(Sample_ID, Site, C2_value, Std)

ploidies_combined <- bind_rows(Oregon_1, Greenhouse)
```



```{r ploidy_plot}
ploidy_plot <- ploidies_combined %>%
  mutate(origin=ifelse(grepl("Oregon",Site), "Field Samples", "Common Garden"),
         Sample_ID.1=ifelse(!grepl("O1", Sample_ID),
                            str_remove_all(Sample_ID,"_.+"), 
                            Sample_ID)) %>% 
  filter(!grepl("Oregon 2", Site)) %>% #group_by(origin) %>% summarize(mean=mean(C2_value), sd=sd(C2_value))
  ggplot(aes(x=Site, y=C2_value))+
  geom_boxplot(outlier.alpha = 0.5, outlier.shape = 16)+ #outlier.colour = "#b94663", outlier.shape = 1, outlier.stroke = 1.5
  geom_point(data= ~subset(., C2_value > 10), stroke=1.5, size=2, pch=1, col="#b94663")+
  theme_classic(base_size = 11)+
  #theme(text = element_text(size=11))+
  #scale_color_manual(values=c("#9750a1","#677ad1"), labels=c("Acer negundo", "Atriplex canescens"))+
  labs(y="C2 value (pg)")+
  facet_wrap(~origin, scales="free_x")+
  geom_text_repel(data= ~subset(., C2_value > 10), aes(label=Sample_ID.1), point.padding = 0.25, nudge_x = .1,
                   nudge_y = 0.5,
                  hjust=0, vjust=0, size=3)

print(ploidy_plot)
ggsave(plot=ploidy_plot, "results/figures/ploidy_plot.pdf", device = "pdf", height = 89, width=89, units = "mm")

```


### Analyze the triploid offspring analyzed with Propidium Iodide (PI)



```{r genome_size_plot}
triploid_offspring_PI.summary <- triploid_offspring_PI %>% 
  filter(grepl("Acru_AS", Std)) %>% 
  group_by(Sample) %>% 
  summarize(mean=mean(C2_Value_ARTR), sd=sd(C2_Value_ARTR)) %>% 
  arrange(mean)
head(triploid_offspring_PI.summary)
write.csv(x=triploid_offspring_PI.summary, "results/tables/genome_sizes_offsprings.csv", row.names = F, quote=F)
```

Plot with common garden samples

```{r genome_sizeplot_all}
tri_plot <- triploid_offspring_PI %>% 
  filter(!grepl("Acne",Std),
       Std=="Acru_AS") %>% 
  mutate(outgr=ifelse(grepl("WAT|IDT",Sample), 
                      "Common Garden",
                      "Field Samples")) %>% 
  mutate(Sample=dplyr::recode(.$Sample,
         IDT_2x = "IDT2_9",
         IDT_3x = "IDT2_5",
         WAT_2x = "WAT1_5",
         WAT_4x = "WAT1_4")) %>% #filter(!grepl("Common", outgr)) %>% dplyr::select(-outgr) %>% 
  ggplot()+
  geom_boxplot(aes(x=reorder(Sample,C2_Value_ARTR), y=C2_Value_ARTR))+
  theme_classic()+
  theme(axis.text.x = element_text(angle = 45, hjust = 1), text = element_text(size=15))+
  labs(x="Sample", y="C2-value (pg)")+
  facet_grid(~ outgr, scales="free_x", space="free_x")
print(tri_plot)
ggsave(plot=tri_plot, "results/figures/tri_plot.pdf", device = "pdf", height = 8, width=8)

```

For the manuscript, we decided to use only the field offspring for the main figure. 

```{r genome_size_plot_fieldsample}
tri_plot_wocg <- triploid_offspring_PI %>% 
  filter(!grepl("Acne",Std),
       Std=="Acru_AS") %>% 
  mutate(outgr=ifelse(grepl("WAT|IDT",Sample), 
                      "Common Garden",
                      "Field Samples")) %>% 
  mutate(Sample=dplyr::recode(.$Sample,
         IDT_2x = "IDT2_9",
         IDT_3x = "IDT2_5",
         WAT_2x = "WAT1_5",
         WAT_4x = "WAT1_4")) %>% filter(!grepl("Common", outgr)) %>%
  dplyr::select(-outgr) %>% 
  ggplot()+
  geom_boxplot(aes(x=reorder(Sample,C2_Value_ARTR), y=C2_Value_ARTR))+
  theme_classic(base_size=15)+
  theme(axis.text.x = element_text(angle = 45, hjust = 1), text = element_text(size=15))+
  labs(x="Sample", y="C2-value (pg)")
print(tri_plot_wocg)
ggsave(plot=tri_plot_wocg, "results/figures/tri_plot_wocg.pdf", device = "pdf", height = 6, width=6)
```

```{r}
sessionInfo()
```




